Cognitive Principles and Individual Differences \\ in Human Syllogistic Reasoning

### 2019-11-11 paper/paper-v3.tex
richards kommentare eingearbeitet

A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in only one of the premises, together with a third middle property that does not appear in the conclusion but appears in both premises. 
\todo[inline]{R: Für mich ist der letzte Satz unverständlich. Vorschlag: ``\dots each of these properties occurs in exactly one of the premises, and there is a third property connecting both premises which does not occur in the conclusion.''}\noindent%

@@

According to~\cite{khemlani:2012}, the majority of participants in experimental studies

@@

Here, we focus on a qualitative assessment of the clustering approach in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.

@@

According to~\cite{declare:2017}, reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
The approach suggested in~\cite{declare:2017} is based on a new computational logic approach for human syllogistic reasoning~\cite{cogsci:2017}, which seems to perform better than other state-of-the-art cognitive theories when evaluated with aggregated data provided in~\cite{khemlani:2012}. 
The approach presented in~\cite{declare:2017} extended~\cite{cogsci:2017} by specifying three reasoning clusters and two heuristic strategies and achieved a match of 92\;\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
Next, the desired\todo{R: desired? Entailed, oder was willst du sagen?} models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
 The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.

@@

\todo[inline]{R: Ist es eine Überlegung wert, \L ukasiewicz logic vor contextual programs einzuführen? Oben benutzen wir schon $\udf$, was erst hier eingeführt wird.}
We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
extended with the $\cxt$ connective, for which the
corresponding truth values are $\top$, $\bot$ and $\udf$, meaning \textit{true}, \textit{false}
and \textit{unknown}, respectively.
% As it seems,  $\cxt$ corresponds to {\L}ukasiewicz necessity operator~\cite{}, where the difference is, that $\cxt$ is restricted to be applicable to literals only.
A \defname{three-valued interpretation}~$I$ is a mapping from the set of
logical formulas to the set
of truth values~$\{ \true, \bot, \udf\}$, represented as a pair~$I~=~\langle I^\true,I^\bot \rangle$ of two disjoint sets of atoms:  $I^\true =  \{A \mid A \mbox{ is mapped to } \true \mbox{ under } I\}$ and $I^\bot = \{A \mid A \mbox{ is mapped to } \bot \mbox{ under } I \}$.
Atoms which do not occur in~$I^\true \cup I^\bot$ are mapped to~$\udf$.
The truth value of a given formula
under $I$ is determined according to the truth tables in
Table~\ref{tab:3vldluka}.
$I(F) = \top$ denotes that a formula $F$ is mapped to true under $I$.
A \defname{three-valued model} $\CalM$ of~$\CalP$ is a three-valued interpretation such that
$\CalM(A \leftarrow \Body) = \true$ for each $A \leftarrow \Body \in \Ground\CalP$.
Let $I = \langle I^\top, I^\bot \rangle$ and $J = \langle J^\top,
J^\bot \rangle$ be two interpretations.
$I \subseteq J$ iff $I^\top \subseteq J^\top$ and $I^\bot \subseteq J^\bot$. $I$ is the \textit{least model} of $\CalP$
iff for any other model~$J$ of $\CalP$ it holds that $I \subseteq J$.

@@

A set of \textit{integrity constraints} $\CalIC$ consists of clauses of
the form $\udf \leftarrow \Body$, where $\Body$ is a conjunction of
literals\todo{R: ``and $\udf$ denotes unknown'' entfernt, sollte an dieser Stelle klar sein}. An interpretation maps
an integrity constraint to $\top$ iff $\Body$ is either mapped to
$\udf$ or $\bot$.  
Given an interpretation~$I$
and a set of integrity constraints~$\CalIC$,~$I$ \textit{satisfies}~$\CalIC$
iff all clauses in~$\CalIC$ are true under~$I$.

@@

The weak completion of contextual
programs might have more than one minimal model, and it is not guaranteed that 
the least fixed point of $\Svlp$ always exists. However,
if these programs do not contain cycles, then there always exists a supported model\todo{R: Begriff ``supported model'' unklar}, which 
corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}.
Let us denote the least fixed point of $\Svlp$ as $\CalM_\CalP$:
As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, we define
$\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
For all programs considered in the sequel, $\CalM_\CalP$ exists.

@@

Eight principles for quantified statements have been developed in~\cite{cogsci:2017,costa:dietz:Hoelldobler:2017,declare:2017}.
In the following, the programs are
specified using the predicates $y$ and $z$, where~$yz$ can be replaced by $ab$, $ba$, $cb$, or $bc$ and depend on the figures shown in Table~\ref{t:2}. First, we introduce the basic principles that are assumed for all premises when applicable. Thus, we assume them to be assumed\todo{R: assumed $\to$ applied?} by all (human) reasoners. 
In Section~\ref{sect:advanced}, we present the advanced principles, which will 
later be used to distinguish between the clusters of reasoners.
% We indicate the cognitive principles in the text by specifying them in brackets.
A complete specification of the logic programs with respect to the moods according to all principles can be found in~\cite{declare:2017}.

@@

\subsubsection{Conditional Representation (\conditionals) and License for Inference (\licenses)}\label{sect:condli}
Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
However, it seems that humans understand them defeasibly, \todo{R: i.e.?} they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008},
i.e.\ they might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
explicitly expressed by 
\textit{all $y$ that are not abnormal are $z$}.
% \\
% Accordingly, $z(X)$ is conjoint with an abnormality predicate: 
% \[\forall X (y(X) \wedge \neg \Ab(X) \rightarrow z(X)).\] 
By default, we assume that \textit{nothing is abnormal}. % with respect to $X$}:
% \[
% \Ab(X) \leftarrow \bot.
% \]

@@

In FOL, the conditional statement $\forall (X) (a(X) \rightarrow b(X))$, is logically
equivalent to its contrapositive, $\forall (X) (\neg b(X) \rightarrow \neg a(X))$.
% This
% contraposition also holds for the syllogistic moods~\MA\ and~\ME. 
There is
evidence that some of the participants might directly or indirectly 
apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994})\todo{R: Die Referenzen gibt es bei mir nicht}.
Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
from Section~\ref{sect:basic}. The clauses that represent the contraposition principles are as follows:
\begin{align}
y'(X) & \leftarrow  \neg z(X) \wedge \neg \Ab_{zy}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \contraposition\ } \\
\Ab_{zy}(X) & \leftarrow \cxt(y(X)). & \tag{ \licenses\ \&\ \contraposition}\\
y(X) & \leftarrow  \neg y'(X) \wedge \neg \Ab_{nyy}(X). & \tag{\transformation\ \&\ \licenses\ \&\ \contraposition} 
\end{align}
The application of (\contraposition) for the \ME\ mood,
\textit{No y is z}, yields \textit{No z is y}, and consists of the following clauses:
\begin{align}
y'(X) & \leftarrow z(X) \wedge \neg \Ab_{zny}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \contraposition\ } \\
\Ab_{zny}(X) & \leftarrow \cxt(y(X)). &\tag{ \licenses\ \&\ \contraposition}\\
y(X) & \leftarrow \neg y'(X) \wedge \neg \Ab_{nyy}(X). & \tag{\transformation\ \&\ \licenses\ \&\ \contraposition} 
% z(o_2) \leftarrow \top. & (\contraposition\  \&\ \import) \\
% \Ab_{nyy}(o_2) \leftarrow \bot. & (\contraposition\  \&\ \licenses\ \&\ \dnegation)
\end{align}

@@

We specify when $\MP$ entails a conclusion with respect to $yz$:\footnote{$yz$ can be replaced by $ac$ or~$ca$.}
\begin{description}
\item[\MA yz (all)] $\CalP \models Ayz$ iff there exists an object $o$ such that $\CalP \ModelsWCS
y(o)$ and for all objects $o$ we find that if $\CalP \ModelsWCS y(o)$ then $\CalP \ModelsWCS z(o)$.
\item [\ME yz (no)]
$\CalP \models Eyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1)$ and for all objects $o_1$ we find that if $\CalP \ModelsWCS y(o_1)$ then
$\CalP \ModelsWCS \neg z(o_1)$.\footnote{
When the contraposition principles applies, then additionally 
the following condition needs to hold as well: There exists an object $o_2$ such that $\CalP \ModelsWCS
z(o_2)$ and for all objects $o_2$ we find that if $\CalP \ModelsWCS z(o_2)$ then
$\CalP \ModelsWCS \neg y(o_2)$.}
\item[\MI yz (some)]
$\CalP \models Iyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1) \wedge z(o_1)$ and there exists an
object $o_2$ such that $\CalP \ModelsWCS y(o_2)$ and
$\CalP \not\ModelsWCS z(o_2)$ and
% there exists an object $o_3$ such that $\CalP \ModelsWCS
% z(o_3) \wedge y(o_3)$ and 
there exists an
object $o_3$ such that $\CalP \ModelsWCS z(o_3)$ and
$\CalP \not\ModelsWCS y(o_3)$.
\item[\MO yz (some are not)]
$\CalP \models Oyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1) \wedge \neg z(o_1)$ and there exists an
object $o_2$ such that $\CalP \ModelsWCS y(o_2)$ and
$\CalP \not\ModelsWCS \neg z(o_2)$.
\item[\NVC\ (no valid conclusion)] iff none of the previous conclusions can be derived.
\end{description}
For \MI yz (some) the conversion is explicitly specified, motivated by the (\converse) principle.
By requiring for \MI yz and \MO yz the existence of an object $o_1$ with the first property $y$,
for which the second property $z$ has to be either either false or unknown,
\todo[inline]{R: Das ist doch nicht richtig, oder? Bei \MI yz muss $z(o_2)$ entweder false oder unknown sein, wie beschrieben, aber bei \MO yz muss es entweder true oder unknown sein.}
\MA yz and \MI yz, (\ME yz and \MO yz, resp.), exclude each other.

@@

Given this sobering outcome with respect to the non-aggregated evaluation, 
we investigated whether patterns could be found on the level of the
cognitive principles: \basic, \contraposition, \abduction\ and \negFailure.
\todo[inline]{R: Ich finde, hier wird nicht ganz klar, was das ``level of cognitive principles'' ist}
Table~\ref{tab:cogprinc} provides an overview of the predicted conclusions according to these principles.\footnote{Note that any advanced principle also includes the logic program clauses specified for \textsf{basic}.}
The predictions of the principles naturally raises a few questions on the plausibility of 
the underlying assumptions:
(i) Do participants use the same reasoning strategy throughout the same reasoning task, i.e.\ 
can their conclusions be 
explained by  the same principles throughout the whole reasoning task, or do they switch strategies within one reasoning task?
(ii) Is the list of principles in Section~\ref{sect:representing} complete, i.e.\ can the conclusions of all reasoners be covered
 by these principles? and
(iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again?
We address these questions in the following.

@@

\cite{khemlani:2016} proposed an approach to model the individual reasoning patterns of the participants, and 
showed that different participants seem to derive different conclusions.
Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory~\cite{johnsonlaird:1983}.
Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
(i) represent possibilities, (ii) principle of iconicity, and (iii) principle of dual processes.
 The procedure of mReasoner consists of two steps:
(1) The construction of a small set of models that represents the terms of the premises, and
(2) the call of a deliberative component to search for counterexamples to conclusions.
They\todo{R: statt ``they'' \cite{khemlani:2016}} specified four parameters for the construction of a mental model and evaluated their
predictions based on 2 datasets, week 1 and week 2, for 20 participants.
The parameters for the construction of the mental models are (i) the 
size of a mental model, (ii) the probability of drawing\todo{R: was heißt draw hier?} the complete set, 
(iii) the probability to search for counter examples,
and (iv) the probability to weaken the conclusion.
The clustering with mReasoner was based on partitioning around\todo{R: around \emph{the} medoids?} medoids clustering algorithm to determine the optimal number of clusters.
The K-means clustering algorithm was used to divide participants into clusters and then
the mReaoner to simulated the clusters by choosing appropriate parameter settings.
\cite{khemlani:2016} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning. However, their approach seems to have two drawbacks. First, some of their 
parameters were of probabilistic nature, i.e.\ there is no explaining factor behind them. Second, the authors did not distinguish between the participants' 
responses in week 1 and week 2, but instead merged the responses. This might give a distorted picture, as participants could
have learned from week 1 to week 2.

>>>>

A conclusion then consists of an assertion about two properties over terms where each of these properties occurs in exactly one of the premises, and there is a third property connecting both premises which does not occur in the conclusion

@@

According to the authors in~\cite{khemlani:2012}, the majority of participants in experimental studies

@@

Here, we focus on a qualitative assessment of the clustering approach proposed in~\cite{declare:2017},
in the sense that we investigate the individual participants' conclusions with respect to the non-aggregated dataset provided by the authors in~\cite{khemlani:2016}.
In Section~\ref{sect:wcsclustering} and~\ref{sect:representing}, we introduce the underlying computational logic approach and the formal representation of cognitive principles within this approach. Section~\ref{sect:aggregateddata} illustrates an example and briefly discusses the performance of the approach presented in~\cite{declare:2017} on non-aggregated data.
Section~\ref{sect:nonag} provides a more adequate characterization of the individuals' reasoning patterns.

@@

Clustering by Principles under WCS first presented in~\cite{declare:2017},
assumes that reasoning clusters can be expressed through cognitive principles and heuristic strategies. Cognitive principles are assumptions (not necessarily valid under CL) made by humans while reasoning, 
whereas heuristic strategies describe the methods that are applied when
conclusions are not derived by reasoning but instead based on structural patterns or {biases}.
This approach is an extension of the computational logic approach for human syllogistic reasoning presented in~\cite{cogsci:2017}: Three reasoning clusters and two heuristic strategies were specified, and achieved a match of 92\;\%, which was the highest result compared to the presented approaches in~\cite{khemlani:2012}.
% The system developed in~\cite{declare:2017}, was extended and participated at the Syllogism Challenge 2017.\footnote{\url{https://www.cc.uni-freiburg.de/modelingchallenge/challenge-2017}} 
Figure~\ref{overview:system:syllchal} illustrates the components of the system:
First, groups of reasoners were specified based on a set of underlying cognitive principles and heuristic strategies.
After that, this set was translated into a logic program (Section~\ref{sect:basic}-\ref{sect:advanced}), representing the group of reasoners.
Next, the respective models were computed (Section~\ref{sect:wcs}-\ref{sect:lfp}).
After that, the entailed relations between $a$ and $c$ were extracted (Section~\ref{sect:entail}). 
 The final answer distribution was specified according to the number of reasoning clusters 
that entailed the conclusion.
Finally, an illustrative representation of these clusters with respect to their distribution was done by means of multinomial process trees~\cite{Riefer1988}.

@@

We consider the three-valued {\L}ukasiewicz logic~\cite{lukasiewicz:20}
extended with the $\cxt$ connective, for which the
corresponding truth values are $\top$, $\bot$ and $\udf$, meaning \textit{true}, \textit{false}
and \textit{unknown}, respectively.
% As it seems,  $\cxt$ corresponds to {\L}ukasiewicz necessity operator~\cite{}, where the difference is, that $\cxt$ is restricted to be applicable to literals only.
A \defname{three-valued interpretation}~$I$ is a mapping from the set of
logical formulas to the set
of truth values~$\{ \true, \bot, \udf\}$, represented as a pair~$I~=~\langle I^\true,I^\bot \rangle$ of two disjoint sets of atoms:  $I^\true =  \{A \mid A \mbox{ is mapped to } \true \mbox{ under } I\}$ and $I^\bot = \{A \mid A \mbox{ is mapped to } \bot \mbox{ under } I \}$.
Atoms which do not occur in~$I^\true \cup I^\bot$ are mapped to~$\udf$.
The truth value of a given formula
under $I$ is determined according to the truth tables in
Table~\ref{tab:3vldluka}.
$I(F) = \top$ denotes that a formula $F$ is mapped to true under $I$.
A \defname{three-valued model} $\CalM$ of~$\CalP$ is a three-valued interpretation such that
$\CalM(A \leftarrow \Body) = \true$ for each $A \leftarrow \Body \in \Ground\CalP$.
Let $I = \langle I^\top, I^\bot \rangle$ and $J = \langle J^\top,
J^\bot \rangle$ be two interpretations.
$I \subseteq J$ iff $I^\top \subseteq J^\top$ and $I^\bot \subseteq J^\bot$. $I$ is the \textit{least model} of $\CalP$
iff for any other model~$J$ of $\CalP$ it holds that $I \subseteq J$.

@@

A set of \textit{integrity constraints} $\CalIC$ consists of clauses of
the form $\udf \leftarrow \Body$, where $\Body$ is a conjunction of
literals. An interpretation maps
an integrity constraint to $\top$ iff $\Body$ is either mapped to
$\udf$ or $\bot$.  
Given an interpretation~$I$
and a set of integrity constraints~$\CalIC$,~$I$ \textit{satisfies}~$\CalIC$
iff all clauses in~$\CalIC$ are true under~$I$.

@@

The weak completion of contextual
programs might have more than one minimal model, and it is not guaranteed that 
the least fixed point of $\Svlp$ always exists. However,
for programs without cycles, there always exists a unique supported model,
which corresponds to the least fixed point of~$\Svlp$~\cite{dietz:hoelldobler:pereira:2017,dietz:hoelldobler:philipp:2017}. This model is \textit{supported} in the sense that 
it will always be computed by $\Svlp$, independent with which interpretation, we start to iterate $\Svlp$ with. Let us denote this model as $\CalM_\CalP$:
% As the fixed point of $\Svlp$ is unique and only computes the supported minimal model of $\WComp\CalP$, 
% we define 
$\CalP \ModelsWCS F$ iff $\CalM_\CalP(F) = \top$.
For all programs considered in the sequel, $\CalM_\CalP$ exists.

@@

Eight principles for quantified statements have been developed in~\cite{cogsci:2017,costa:dietz:Hoelldobler:2017,declare:2017}.
In the following, the programs are
specified using the predicates $y$ and $z$, where~$yz$ can be replaced by $ab$, $ba$, $cb$, or $bc$ and depend on the figures shown in Table~\ref{t:2}. First, we introduce the basic principles that are assumed for all premises when applicable.
In Section~\ref{sect:advanced}, we present the advanced principles, which will 
later be used to distinguish between the clusters of reasoners.
% We indicate the cognitive principles in the text by specifying them in brackets.
A complete specification of the logic programs with respect to the moods according to all principles can be found in~\cite{declare:2017}.

@@

\subsubsection{Conditional Representation (\conditionals) and License for Inference (\licenses)}\label{sect:condli}
Consider \textit{All y are z} ({\MA yz}), which in CL, is represented as $\forall X (y(X) \rightarrow z(X))$.
Similarly, we assume that humans also understand this statement as a conditional sentence (\conditionals).
However, it seems that humans understand them defeasibly, i.e., they assume a \textit{license for the inference}~\cite{stenning:vanlambalgen:2008} and might allow exceptions to the conditional (\licenses). For (\MA yz) this can be 
explicitly expressed by 
\textit{all $y$ that are not abnormal are $z$}.
% \\
% Accordingly, $z(X)$ is conjoint with an abnormality predicate: 
% \[\forall X (y(X) \wedge \neg \Ab(X) \rightarrow z(X)).\] 
By default, we assume that \textit{nothing is abnormal}. % with respect to $X$}:
% \[
% \Ab(X) \leftarrow \bot.
% \]

@@

In FOL, the conditional statement $\forall (X) (a(X) \rightarrow b(X))$, is logically
equivalent to its contrapositive, $\forall (X) (\neg b(X) \rightarrow \neg a(X))$.
% This
% contraposition also holds for the syllogistic moods~\MA\ and~\ME. 
There is
evidence that some of the participants might directly or indirectly 
apply the contrapositive (\contraposition) within a reasoning task (cf.~\cite{rips:1994,mentallogic:1994}).
Consider again $\CalP_{Ayz}$, the logic program representing (\MA yz)
from Section~\ref{sect:basic}. The clauses that represent the contraposition principles are as follows:
\begin{align}
y'(X) & \leftarrow  \neg z(X) \wedge \neg \Ab_{zy}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \contraposition\ } \\
\Ab_{zy}(X) & \leftarrow \cxt(y(X)). & \tag{ \licenses\ \&\ \contraposition}\\
y(X) & \leftarrow  \neg y'(X) \wedge \neg \Ab_{nyy}(X). & \tag{\transformation\ \&\ \licenses\ \&\ \contraposition} 
\end{align}
The application of (\contraposition) for the \ME\ mood,
\textit{No y is z}, yields \textit{No z is y}, and consists of the following clauses:
\begin{align}
y'(X) & \leftarrow z(X) \wedge \neg \Ab_{zny}(X). & \tag{\conditionals\ \&\ \licenses\ \&\ \contraposition\ } \\
\Ab_{zny}(X) & \leftarrow \cxt(y(X)). &\tag{ \licenses\ \&\ \contraposition}\\
y(X) & \leftarrow \neg y'(X) \wedge \neg \Ab_{nyy}(X). & \tag{\transformation\ \&\ \licenses\ \&\ \contraposition} 
% z(o_2) \leftarrow \top. & (\contraposition\  \&\ \import) \\
% \Ab_{nyy}(o_2) \leftarrow \bot. & (\contraposition\  \&\ \licenses\ \&\ \dnegation)
\end{align}

@@

We specify when $\MP$ entails a conclusion with respect to $yz$:\footnote{$yz$ can be replaced by $ac$ or~$ca$.}
\begin{description}
\item[\MA yz (all)] $\CalP \models Ayz$ iff there exists an object $o$ such that $\CalP \ModelsWCS
y(o)$ and for all objects $o$ we find that if $\CalP \ModelsWCS y(o)$ then $\CalP \ModelsWCS z(o)$.
\item [\ME yz (no)]
$\CalP \models Eyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1)$ and for all objects $o_1$ we find that if $\CalP \ModelsWCS y(o_1)$ then
$\CalP \ModelsWCS \neg z(o_1)$.\footnote{
When the contraposition principles applies, then additionally 
the following condition needs to hold as well: There exists an object $o_2$ such that $\CalP \ModelsWCS
z(o_2)$ and for all objects $o_2$ we find that if $\CalP \ModelsWCS z(o_2)$ then
$\CalP \ModelsWCS \neg y(o_2)$.}
\item[\MI yz (some)]
$\CalP \models Iyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1) \wedge z(o_1)$ and there exists an
object $o_2$ such that $\CalP \ModelsWCS y(o_2)$ and
$\CalP \not\ModelsWCS z(o_2)$ and
% there exists an object $o_3$ such that $\CalP \ModelsWCS
% z(o_3) \wedge y(o_3)$ and 
there exists an
object $o_3$ such that $\CalP \ModelsWCS z(o_3)$ and
$\CalP \not\ModelsWCS y(o_3)$.
\item[\MO yz (some are not)]
$\CalP \models Oyz$ iff there exists an object $o_1$ such that $\CalP \ModelsWCS
y(o_1) \wedge \neg z(o_1)$ and there exists an
object $o_2$ such that $\CalP \ModelsWCS y(o_2)$ and
$\CalP \not\ModelsWCS \neg z(o_2)$.
\item[\NVC\ (no valid conclusion)] iff none of the previous conclusions can be derived.
\end{description}
For \MI yz (some) the conversion is explicitly specified, motivated by the (\converse) principle.
By requiring for \MI yz (\MO yz, resp) the existence of an object with the first property $y$,
for which the second property $z$ has to be either either false or unknown (true or unknown, resp.),
\MA yz and \MI yz, (\ME yz and \MO yz, resp.), exclude each other.

@@

Given this sobering outcome with respect to the non-aggregated evaluation, 
we ignored the clusters, and investigated whether parallels between the applied cognitive principles could be observed: \basic, \contraposition, \abduction\ and \negFailure.
Table~\ref{tab:cogprinc} provides an overview of the predicted conclusions according to these principles.\footnote{Note that any advanced principle also includes the logic program clauses specified for \textsf{basic}.}
The predictions of the principles naturally raises a few questions on the plausibility of 
the underlying assumptions:
(i) Do participants use the same reasoning strategy throughout the same reasoning task, i.e.\ 
can their conclusions be 
explained by  the same principles throughout the whole reasoning task, or do they switch strategies within one reasoning task?
(ii) Is the list of principles in Section~\ref{sect:representing} complete, i.e.\ can the conclusions of all reasoners be covered
 by these principles? and
(iii) Do participants learn, i.e.\ do they improve their logical ability by doing the same reasoning task again?
We address these questions in the following.

@@

\cite{khemlani:2016} proposed an approach to model the individual reasoning patterns of the participants, and 
showed that different participants seem to derive different conclusions.
Their approach is built on top of the mReasoner~\cite{khemlani:2013}, an extension of the mental model theory~\cite{johnsonlaird:1983}.
Reasoning depends on the construction and manipulation of mental models, based on 3 fundamental principles: 
(i) represent possibilities, (ii) principle of iconicity, and (iii) principle of dual processes.
 The procedure of mReasoner consists of two steps:
(1) The construction of a small set of models that represents the terms of the premises, and
(2) the call of a deliberative component to search for counterexamples to conclusions.
They specified four parameters for the construction of a mental model and evaluated their
predictions based on 2 datasets, week 1 and week 2, for 20 participants.
The parameters for the construction of the mental models are (i) the 
size of a mental model, (ii) the probability of drawing a conclusions only from the most common mental models, or from the complete set,
(iii) the probability to search for counter examples,
and (iv) the probability to weaken the conclusion.
The clustering with mReasoner was based on a particular clustering algorithm to determine the optimal number of clusters for their dataset.
The K-means clustering algorithm was used to divide participants into clusters and then
the mReaoner to simulated the clusters by choosing appropriate parameter settings.
\cite{khemlani:2016} identified three different reasoning clusters: The intuitive cluster, the intermediate cluster and the deliberative cluster.
The approach is novel in the sense that until now no other cognitive theory tried to rigorously account for individual differences in human syllogistic reasoning. However, their approach seems to have two drawbacks. First, some of their 
parameters were of probabilistic nature, i.e.\ there is no explaining factor behind them. Second, the authors did not distinguish between the participants' 
responses in week 1 and week 2, but instead merged the responses. This might give a distorted picture, as participants could
have learned from week 1 to week 2.
